---
title: "Case report workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case report workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows a more realistic case-investigation workflow with uncertain exposures, sick contacts, and export-ready output.

## Example reports

```{r}
library(epistract)

reports <- tibble::tibble(
  case_id = 1:3,
  text = c(
    paste(
      "Case 1: Priya Sharma.",
      "Test Result: Positive for Salmonella.",
      "Case notified via laboratory report (Dorevitch Pathology, Heidelberg) on 14 March 2024.",
      "Initial interview conducted 15 March.",
      "The case reported onset of diarrhoea and abdominal cramping late evening of 12 March.",
      "Takeaway chicken wrap with mayonnaise from Cluck & Go on Swan Street, Richmond on 11 March.",
      "Also attended a family barbecue on 10 March at Fawkner Park, South Yarra.",
      "No overseas travel reported.",
      "Household contact (partner) reported mild gastrointestinal symptoms but was not tested."
    ),
    paste(
      "Case 2: Liam O'Connor.",
      "Test Result: Positive for Salmonella.",
      "Notified 16 March 2024 following positive stool culture.",
      "Interview completed same day.",
      "The case reported eating at The Rusty Bean on Sydney Road, Brunswick on 13 March.",
      "Also consumed pre-packaged spinach and feta salad from Woolworths Barkly Square earlier that week.",
      "No known sick contacts.",
      "Works in an office setting at TechHub CBD."
    ),
    paste(
      "Case 3: Mei Lin Chen.",
      "Test Result: Positive for Salmonella.",
      "Lab notification received 13 March.",
      "Case interview delayed until 17 March due to inability to contact.",
      "Symptoms included diarrhoea, fatigue, and abdominal discomfort.",
      "Exposure history included attendance at the Coburg Night Market on 9 March.",
      "One housemate reportedly unwell but not confirmed."
    )
  )
)
```

## Run extraction

```{r, eval = FALSE}
llm <- llm()

out <- extract_epi_data(
  reports,
  input_col = text,
  type = type_epi_case_report(),
  llm = llm
)
```

## Working with repeated fields

The output can include list-columns for repeated exposures and contacts.

```{r}
repeated_fields <- c("exposures", "contacts", "illness.symptoms")
repeated_fields
```

Those list-columns are useful during analysis, but they cannot be written directly with `write.csv()`.

## Export to CSV or TSV

```{r}
example_out <- tibble::tibble(
  case_id = 1L,
  illness.symptoms = list(c("diarrhoea", "fever")),
  contacts = list(tibble::tibble(
    relationship = c("partner", "housemate"),
    tested = c("no", "unknown")
  ))
)

prepare_epi_export(example_out)
```

```{r, eval = FALSE}
export_ready <- prepare_epi_export(out)
write.csv(export_ready, "example.csv", row.names = FALSE)

write_epi_delim(out, "example.tsv", sep = "\t")
```

## When to use this workflow

Use this pattern when you have:

- one report per row
- a mostly stable schema across all reports
- repeated exposures or contacts that should be preserved for later analysis
