---
title: "Custom schemas and prompts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom schemas and prompts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette covers more advanced usage: simplifying the schema, narrowing the prompt, and targeting a fit-for-purpose export for a specific surveillance task.

## Why customize?

Local models often perform better when the requested schema is smaller and more specific.

For example, if you only need disease, onset, exposures, and contacts, a narrower schema can improve extraction consistency.

## Build a smaller schema

```{r}
library(epistract)

mini_case_type <- ellmer::type_object(
  "Minimal enteric disease extraction schema.",
  case_name = ellmer::type_string("Case name as reported.", required = FALSE),
  disease = ellmer::type_string("Disease named in the report.", required = FALSE),
  onset_date = ellmer::type_string("Onset date in YYYY-MM-DD when possible.", required = FALSE),
  symptoms = ellmer::type_array(
    ellmer::type_string("One symptom."),
    description = "Reported symptoms.",
    required = FALSE
  ),
  exposures = ellmer::type_array(
    ellmer::type_string("One possible exposure."),
    description = "Possible food, venue, or event exposures.",
    required = FALSE
  ),
  contacts = ellmer::type_array(
    ellmer::type_string("One relevant contact."),
    description = "Relevant household or close contacts.",
    required = FALSE
  )
)

mini_case_type
```

## Use a targeted prompt

```{r}
mini_prompt <- function(text) {
  paste(
    "Extract only the case name, disease, onset, symptoms, exposures, and contacts.",
    "Do not include any extra fields.",
    "Use missing values when the text is unclear.",
    text,
    sep = "\n\n"
  )
}
```

## Run the narrower extraction

```{r, eval = FALSE}
llm <- llm(model = "gemma3:4b")

mini_out <- extract_epi_data(
  reports,
  input_col = text,
  type = mini_case_type,
  llm = llm,
  prompt = mini_prompt
)
```

## Export only the columns you need

```{r}
mini_export <- tibble::tibble(
  case_name = "Priya Sharma",
  symptoms = list(c("diarrhoea", "abdominal cramping")),
  exposures = list(c("Cluck & Go chicken wrap", "family barbecue"))
)

prepare_epi_export(mini_export)
```

## Summary

Increase complexity gradually:

1. start with the default `type_epi_case_report()` schema
2. inspect the extracted columns and list-columns
3. if needed, reduce the schema and tighten the prompt
4. export with `prepare_epi_export()` or `write_epi_delim()`
