---
title: "Getting started with epistract"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with epistract}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`epistract` extracts structured epidemiological fields from free-text case reports using `ellmer` structured outputs and a local Ollama model.

This first vignette shows the smallest useful workflow:

1. create a small table of reports
2. create an LLM client with `llm()`
3. extract structured fields with `extract_epi_data()`
4. inspect the flattened output

## Minimal input

```{r}
library(epistract)

reports <- tibble::tibble(
  case_id = 1:2,
  text = c(
    paste(
      "Case 1: Priya Sharma.",
      "Test Result: Positive for Salmonella.",
      "Case notified via laboratory report on 14 March 2024.",
      "Initial interview conducted 15 March.",
      "Onset of diarrhoea and abdominal cramping late evening of 12 March."
    ),
    paste(
      "Case 2: Liam O'Connor.",
      "Positive stool culture for Salmonella notified 16 March 2024.",
      "Interview completed same day.",
      "Onset sometime early on the 14th with nausea, vomiting, and later diarrhoea."
    )
  )
)

reports
```

## Create the model client

```{r, eval = FALSE}
llm <- llm(model = "gemma3:4b")
```

## Extract structured fields

```{r, eval = FALSE}
out <- extract_epi_data(
  reports,
  input_col = text,
  type = type_epi_case_report(),
  llm = llm
)
```

The default schema is already tuned for single case investigation records, so a basic call often works well as a starting point.

## Expected output shape

The extraction result is a flat tibble with dot-notated columns for nested fields.

```{r}
example_names <- c(
  "case_id",
  "text",
  "case.case_status",
  "case.patient.full_name",
  "notification.disease",
  "notification.test_result",
  "interview.interview_date",
  "illness.onset_date",
  "illness.symptoms"
)

example_names
```

## Next steps

- Use `type_epi_case_report()` for case interview narratives.
- Use `prepare_epi_export()` before writing CSV files.
- Move to the next vignette for more realistic exposure histories and repeated nested fields.
